Pre- and Post-Debate Democratic Primary Data: Twitter, Google Trends, and Polls

### 2019-12-12 Paper/paper-draft-Padgett-Thiam.Rmd
knit new version

```{r, message="FALSE", warning=FALSE, results="hide", include=FALSE}
opts_chunk$set(tidy.opts=list(width=60),tidy=TRUE)
```
\newpage
## Introduction

@@

```{r, eval=FALSE, tidy=TRUE}

@@

```{r, eval=FALSE, tidy=TRUE}
tweets$BS <-(str_detect(tweets$text, 
                        regex("#Sanders|@BernieSanders", 
                              ignore_case = TRUE)))
tweets$KH <-(str_detect(tweets$text, 
                        regex("#harris |@KamalaHarris",  
                              ignore_case = TRUE)))
tweets$JB <-(str_detect(tweets$text, 
                        regex("#biden |@JoeBiden", 
                              ignore_case = TRUE)))
tweets$EW <-(str_detect(tweets$text, 
                        regex("#warren |@ewarren", 
                              ignore_case = TRUE)))
tweets$PB <-(str_detect(tweets$text, 
                        regex("#buttigieg|@PeteButtigieg", 
                              ignore_case = TRUE)))
tweets$TS <-(str_detect(tweets$text, 
                        regex("#steyer | @TomSteyer", 
                              ignore_case = TRUE)))
tweets$AY <-(str_detect(tweets$text, 
                        regex("#yang| @AndrewYang", 
                              ignore_case = TRUE)))
tweets$BC <-(str_detect(tweets$text, 
                        regex("#booker | @BookerCory", 
                              ignore_case = TRUE)))
tweets$AK <-(str_detect(tweets$text, 
                        regex("#klobuchar | @amyklobuchar", 
                              ignore_case = TRUE)))
tweets$TG <-(str_detect(tweets$text, 
                        regex("#gabbard |@TulsiGabbard", 
                              ignore_case = TRUE)))
```

@@

All the tweets received before 9 p.m. on November 20th were accounted for in the pre-debate dataset and all tweets from 11 p.m. on November 20th to the next day are in the post-debate dataset. Note that tweets during the debate (9 - 11 p.m.) were removed.
```{r, eval=FALSE, tidy=TRUE}
pre_debate <- tweets %>%
              filter(date(date) == "2019-11-20" & hour(date) < 21 )

@@

```{r, eval=FALSE, tidy=TRUE}
pre_sent_pos <- data.frame("Candidate" = c("Biden", 
                                           "Sanders", 
                                           "Warren", 
                                           "Harris", 
                                           "Buttigieg"), 
 "Positive" =c((sum(sentiments_jb$PositivityGI))/27477,
(sum(sentiments_bs$PositivityGI))/31603,                                                                    (sum(sentiments_ew$PositivityGI))/18628,
(sum(sentiments_kh$PositivityGI))/16362,
(sum(sentiments_pb$PositivityGI))/27242))
```

@@

```{r, eval=FALSE, tidy=TRUE}
register_google(key = "XXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXX")
res1 <- gtrends(c("Joe Biden", "Bernie Sanders", 
                  "Elizabeth Warren", "Kamala Harris", 
                  "Pete Buttigieg"), geo = "US", 
time = "2019-11-18 2019-11-22", low_search_volume = T)

@@

```{r, tidy=TRUE}
github_link <- "https://github.com/znpadgett/surv727_padgett_thiam/raw/master/Data/Project%20polling%20data.xlsx"
temp_file <- tempfile(fileext = ".xlsx")
req <- GET(github_link, 
           write_disk(path = temp_file))
polling_data <- readxl::read_excel(temp_file)
```

@@

```{r, tidy=TRUE, fig.cap="Pre- and Post-debate Google Trends Data, by Candidate"}
interest_over_time_all$Pre_post <- factor(interest_over_time_all$Pre_post, 
                                          levels = c("Pre-debate", "Post-debate"))

@@

For the polling data, we looked at the pre- and post-debate polling numbers for each candidate.
```{r, tidy=TRUE, fig.cap="Pre- and Post-debate Polling Data, by Candidate"}
polling_data$Pre_post <- factor(polling_data$Pre_post, 
levels = c("Pre-debate", "Post-debate"))

@@

```{r, tidy=TRUE, warning=FALSE, message=FALSE}
#clean data
gtrends <-
  interest_over_time_all %>%
  mutate(Candidate=keyword, Data="GTrends") %>%
  filter(Candidate %in% c("Joe Biden", "Bernie Sanders", 
                          "Elizabeth Warren", "Kamala Harris", 
                          "Pete Buttigieg")) %>%
  group_by(Candidate, Pre_post, Data) %>%
  arrange(Candidate, Pre_post) %>%
  summarise(Percentage=mean(hits))

@@

Then, we created a shiny app, which can be viewed at https://znpadgett.shinyapps.io/surv727app/:
```{r, eval=FALSE, tidy=TRUE}
# Define UI

@@

```{r, tidy=TRUE, fig.cap="Comparison of Post-debate Twitter and Polling Data"}
graphing_data <- cbind(twitter, poll, gtrends)
graphing_data %>%
  group_by(Pre_post, Candidate) %>%
  filter(Pre_post=="Post-debate") %>%
  ggplot() +
  geom_point(mapping = aes(x=Percentage, y=Percentage1, 
                           color=Candidate), size=4) +
  xlim(0,20) +
  ylim(0,30) +
  labs(x="Twitter Sentiment", 
       y="Polling Data")
```

@@

```{r, tidy=TRUE, fig.cap="Comparison of Post-debate Google Trends and Polling Data"}
graphing_data  %>%
  group_by(Pre_post, Candidate) %>%
  filter(Pre_post=="Post-debate") %>%
  ggplot() +
  geom_point(mapping = aes(x=Percentage2, y=Percentage1, 
                           color=Candidate), size=4) +
  xlim(0,100) +
  ylim(0,30) +
  labs(x="Google Trends Data", 
       y="Polling Data")
```

@@

```{r, tidy=TRUE, fig.cap="Comparison of Changes in Twitter and Polling Data"}
pre <-
graphing_data  %>%
  select(Pre_post, Candidate, Percentage, Percentage1, Percentage2) %>%
  group_by(Pre_post, Candidate, Data) %>%
  filter(Pre_post=="Pre-debate") %>%
  mutate(Pre_twitter=Percentage, Pre_polling=Percentage1, 
         Pre_gtrends=Percentage2)

@@

```{r, tidy=TRUE, fig.cap="Comparison of Changes in Google Trends and Polling Data"}
change_data %>%
  group_by(Pre_post, Candidate) %>%
  mutate(change_twitter = ((Post_twitter - Pre_twitter) / 
                             Pre_twitter) * 100, 
         change_polling = ((Post_polling - Pre_polling) / 
                             Pre_polling) * 100, 
         change_gtrends = ((Post_gtrends - Pre_gtrends) / 
                             Pre_gtrends) * 100) %>%
  ggplot() +
  geom_point(mapping = aes(x=change_gtrends, y=change_polling, 
                           color=Candidate), size=4) +
  labs(x="Percent Change in Google Trends Data", 
       y="Percent Change in Polling Data")
```

>>>>

\newpage
## Introduction

@@

```{r, eval=FALSE}

@@

```{r, eval=FALSE}
tweets$BS <-(str_detect(tweets$text, regex("#Sanders|@BernieSanders", 
                              ignore_case = TRUE)))
tweets$KH <-(str_detect(tweets$text, regex("#harris |@KamalaHarris",  
                              ignore_case = TRUE)))
tweets$JB <-(str_detect(tweets$text, regex("#biden |@JoeBiden", 
                              ignore_case = TRUE)))
tweets$EW <-(str_detect(tweets$text, regex("#warren |@ewarren", 
                              ignore_case = TRUE)))
tweets$PB <-(str_detect(tweets$text, regex("#buttigieg|@PeteButtigieg", 
                              ignore_case = TRUE)))
tweets$TS <-(str_detect(tweets$text, regex("#steyer | @TomSteyer", 
                              ignore_case = TRUE)))
tweets$AY <-(str_detect(tweets$text, regex("#yang| @AndrewYang", 
                              ignore_case = TRUE)))
tweets$BC <-(str_detect(tweets$text, regex("#booker | @BookerCory", 
                              ignore_case = TRUE)))
tweets$AK <-(str_detect(tweets$text, regex("#klobuchar | @amyklobuchar", 
                              ignore_case = TRUE)))
tweets$TG <-(str_detect(tweets$text, regex("#gabbard |@TulsiGabbard", 
                              ignore_case = TRUE)))
```

@@

All the tweets received before 9 p.m. on November 20th were accounted for in the pre-debate dataset and all tweets from 11 p.m. on November 20th to the next day are in the post-debate dataset. Note that tweets during the debate (9 - 11 p.m.) were removed.
```{r, eval=FALSE}
pre_debate <- tweets %>%
              filter(date(date) == "2019-11-20" & hour(date) < 21 )

@@

```{r, eval=FALSE}
pre_sent_pos <- data.frame("Candidate" = c("Biden", 
                                           "Sanders", 
                                           "Warren", 
                                           "Harris", 
                                           "Buttigieg"), 
 "Positive" =c((sum(sentiments_jb$PositivityGI))/27477,
(sum(sentiments_bs$PositivityGI))/31603,                                                                    (sum(sentiments_ew$PositivityGI))/18628,
(sum(sentiments_kh$PositivityGI))/16362,
(sum(sentiments_pb$PositivityGI))/27242))
```

@@

```{r, eval=FALSE}
register_google(key = "XXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXX")
res1 <- gtrends(c("Joe Biden", "Bernie Sanders", 
                  "Elizabeth Warren", "Kamala Harris", 
                  "Pete Buttigieg"), geo = "US", 
time = "2019-11-18 2019-11-22", low_search_volume = T)

@@

```{r}
github_link <- "https://github.com/znpadgett/surv727_padgett_thiam/raw/master/Data/Project%20polling%20data.xlsx"
temp_file <- tempfile(fileext = ".xlsx")
req <- GET(github_link, 
           write_disk(path = temp_file))
polling_data <- readxl::read_excel(temp_file)
```

@@

```{r, fig.cap="Pre- and Post-debate Google Trends Data, by Candidate"}
interest_over_time_all$Pre_post <- factor(interest_over_time_all$Pre_post, 
                                          levels = c("Pre-debate", "Post-debate"))

@@

For the polling data, we looked at the pre- and post-debate polling numbers for each candidate.
```{r, fig.cap="Pre- and Post-debate Polling Data, by Candidate"}
polling_data$Pre_post <- factor(polling_data$Pre_post, 
levels = c("Pre-debate", "Post-debate"))

@@

```{r, warning=FALSE, message=FALSE}
#clean data
gtrends <-
  interest_over_time_all %>%
  mutate(Candidate=keyword, Data="GTrends") %>%
  filter(Candidate %in% c("Joe Biden", "Bernie Sanders", 
                          "Elizabeth Warren", "Kamala Harris", 
                          "Pete Buttigieg")) %>%
  group_by(Candidate, Pre_post, Data) %>%
  arrange(Candidate, Pre_post) %>%
  summarise(Percentage=mean(hits))

@@

Then, we created a shiny app, which can be viewed at https://znpadgett.shinyapps.io/surv727app/:
```{r, eval=FALSE}
# Define UI

@@

```{r, fig.cap="Comparison of Post-debate Twitter and Polling Data"}
graphing_data <- cbind(twitter, poll, gtrends)
graphing_data %>%
  group_by(Pre_post, Candidate) %>%
  filter(Pre_post=="Post-debate") %>%
  ggplot() +
  geom_point(mapping = aes(x=Percentage, y=Percentage1, 
                           color=Candidate), size=4) +
  xlim(0,20) +
  ylim(0,30) +
  labs(x="Twitter Sentiment", 
       y="Polling Data")
```

@@

```{r, fig.cap="Comparison of Post-debate Google Trends and Polling Data"}
graphing_data  %>%
  group_by(Pre_post, Candidate) %>%
  filter(Pre_post=="Post-debate") %>%
  ggplot() +
  geom_point(mapping = aes(x=Percentage2, y=Percentage1, 
                           color=Candidate), size=4) +
  xlim(0,100) +
  ylim(0,30) +
  labs(x="Google Trends Data", 
       y="Polling Data")
```

@@

```{r, fig.cap="Comparison of Changes in Twitter and Polling Data"}
pre <-
graphing_data  %>%
  select(Pre_post, Candidate, Percentage, Percentage1, Percentage2) %>%
  group_by(Pre_post, Candidate, Data) %>%
  filter(Pre_post=="Pre-debate") %>%
  mutate(Pre_twitter=Percentage, Pre_polling=Percentage1, 
         Pre_gtrends=Percentage2)

@@

```{r, fig.cap="Comparison of Changes in Google Trends and Polling Data"}
change_data %>%
  group_by(Pre_post, Candidate) %>%
  mutate(change_twitter = ((Post_twitter - Pre_twitter) / 
                             Pre_twitter) * 100, 
         change_polling = ((Post_polling - Pre_polling) / 
                             Pre_polling) * 100, 
         change_gtrends = ((Post_gtrends - Pre_gtrends) / 
                             Pre_gtrends) * 100) %>%
  ggplot() +
  geom_point(mapping = aes(x=change_gtrends, y=change_polling, 
                           color=Candidate), size=4) +
  labs(x="Percent Change in Google Trends Data", 
       y="Percent Change in Polling Data")
```
